---
title: "Report: Predicted variants and effects"
author: "`r sessionInfo()$platform`"
date: "`r format(Sys.time(), '%d %B, %Y')`"
params:
  rmd: "report.Rmd"
output:
  html_document:
    theme: cosmo
    toc: yes
    toc_depth: 2
    number_sections: yes
    df_print: paged
---

----------

# Background

This report summarizes the genetic variants and their predicted effects from running the [`snakemake-simple-mapping`](https://github.com/MPUSP/snakemake-simple-mapping) workflow.

----------

# Prerequisites

## Packages

Loaded required R packages:

- `tidyverse`
- `ggrepel`
- `ggpubr`
- `ggupset`
- `GGally`
- `GenomicRanges`
- `GenomicFeatures`
- `Biostrings`


```{r}
suppressPackageStartupMessages({
  library(tidyverse)
  library(ggrepel)
  library(ggpubr)
  library(ggupset)
  library(GenomicRanges)
  library(GenomicFeatures)
  library(Biostrings)
})
```

```{r}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r}
# messages
messages <- c()

# import required parameters
sm_params <- snakemake@params
vcf_count <- sm_params$vcf_count
figure_dir <- gsub("report.html$", "figures/", snakemake@output$html)
if (!dir.exists(figure_dir)) {
  dir.create(figure_dir)
  messages <- append(messages, paste0("created figure dir: ", figure_dir))
}

# custom ggplot2 theme that is reused for all later plots
custom_colors <- c(
  "#E7298A", "#66A61E", "#E6AB02", "#7570B3",
  "#B3B3B3", "#1B9E77", "#D95F02", "#A6761D"
)

custom_range <- function(n = 5, quantitative = FALSE) {
  if (quantitative) {
    colorRampPalette(custom_colors[c(5, 3, 4)])(n)
  } else {
    colorRampPalette(custom_colors)(n)
  }
}

custom_theme <- function(
    base_size = 12, base_line_size = 1.0,
    base_rect_size = 1.0, ...) {
  theme_light(
    base_size = base_size,
    base_line_size = base_line_size,
    base_rect_size = base_rect_size
  ) +
    theme(
      title = element_text(colour = grey(0.4), size = 10),
      plot.margin = unit(c(12, 12, 12, 12), "points"),
      axis.ticks.length = unit(0.2, "cm"),
      axis.ticks = element_line(
        colour = grey(0.4),
        linetype = "solid", lineend = "round"
      ),
      axis.text.x = element_text(colour = grey(0.4), size = 10),
      axis.text.y = element_text(colour = grey(0.4), size = 10),
      panel.grid.major = element_line(
        linewidth = 0.6,
        linetype = "solid", colour = grey(0.9)
      ),
      panel.grid.minor = element_blank(),
      panel.border = element_rect(
        linetype = "solid",
        colour = grey(0.4), fill = NA, linewidth = 1.0
      ),
      panel.background = element_blank(),
      strip.background = element_blank(),
      strip.text = element_text(
        colour = grey(0.4), size =
          10, margin = unit(rep(3, 4), "points")
      ),
      legend.text = element_text(colour = grey(0.4), size = 10),
      legend.title = element_blank(),
      legend.background = element_blank(),
      ...
    )
}

# set graphical parameter for subfigure labels
list_fontpars <- list(face = "plain", size = 14)

# default figure size
figwidth <- 9.5
figheight <- 4.5

# function to export an image as svg and png
save_plot <- function(
    pl, path = figure_dir,
    width = figwidth, height = figheight, index = NA) {
  pl_name <- deparse(substitute(pl))
  if (!is.na(index)) pl_name <- paste0(pl_name, "_", index)
  svg(
    filename = paste0(path, pl_name, ".svg"),
    width = width, height = height
  )
  print(pl)
  dev.off()
  png(
    filename = paste0(path, pl_name, ".png"),
    width = width * 125, height = height * 125, res = 120
  )
  print(pl)
  invisible(capture.output(dev.off()))
}
```

## Configuration

All options that were used to run this workflow are listed at [the end of this report](#about-this-report).

## Variant data

Variant data was imported as `.vcf` files. The primary output from the workflow that is used 
here are the variant calling files including the variant effect prediction, one for each sample.

```{r}
# import variant files
list_variants <- unlist(snakemake@input$variants)
print(list_variants)
```

The imported variant data has the following layout (first 10 rows shown):

- *NOTE: tables are only rendered in HTML output but not PDF.*

```{r}
df <- lapply(list_variants, function(file) {
  properties <- strsplit(file, "/")[[1]]
  read_tsv(file, show_col_types = FALSE, comment = "##") %>%
    dplyr::rename(CHROM = `#CHROM`) %>%
    dplyr::select(1:8) %>%
    mutate(across(all_of(c("POS", "QUAL")), as.numeric)) %>%
    mutate(INFO_split = str_split(INFO, ";")) %>%
    mutate(INFO_keys = map(INFO_split, ~ str_extract(.x, "^[^=]+"))) %>%
    mutate(INFO_values = map(INFO_split, ~ str_remove(.x, "^[^=]+="))) %>%
    mutate(INFO_df = map2(INFO_keys, INFO_values, ~ set_names(as.list(.y), .x))) %>%
    unnest_wider(INFO_df) %>%
    dplyr::select(-INFO, -INFO_split, -INFO_keys, -INFO_values)
}) %>%
  set_names(basename(list_variants)) %>%
  bind_rows(.id = "FILE") %>%
  mutate(FILE = gsub("_(vep|snpeff)\\.vcf\\.gz", "", FILE))

# parse the variant effect column
if ("CSQ" %in% colnames(df)) {
  df <- df %>%
    mutate(CSQ = str_split(CSQ, "\\|")) %>%
    unnest_wider(CSQ, names_sep = "_") %>%
    dplyr::rename(
      VARTYPE = `CSQ_2`, IMPACT = `CSQ_3`, GENE = `CSQ_4`,
      AA_CHANGE = `CSQ_11`, CODON_CHANGE = `CSQ_16`
    ) %>%
    dplyr::select(-matches("CSQ_")) %>%
    mutate(VARTYPE = gsub("coding_sequence_variant&", "", VARTYPE))
}

head(df, n = 10)
```

```{r}
# import consensus file
list_consensus <- snakemake@input$consensus
df_consensus <- read_tsv(
  list_consensus,
  show_col_types = FALSE,
  col_names = c("CHROM", "POS", "REF", "ALT", "APPEARANCE")
) %>%
  mutate(POS = as.numeric(POS))
```

```{r}
# import genome fasta
input_fasta <- snakemake@input$fasta
genome_dna <- readDNAStringSet(input_fasta)
seqlevels(genome_dna) <- str_split_i(seqlevels(genome_dna), " ", i = 1)
messages <- append(messages, "imported genome fasta file")

# import GFF for genome
quiet_txdb <- quietly(GenomicFeatures::makeTxDbFromGFF)
txdb_msg <- quiet_txdb(file = snakemake@input$gff)
txdb <- txdb_msg$result
annotated_transcripts <- transcripts(txdb)
messages <- append(messages, paste0("warning: ", txdb_msg$warnings))
messages <- append(messages, txdb_msg$messages)
```

## Variant analysis

### Variant frequency by sample {.tabset}

- the following section is an overview about variants per sample
- the first figure shows the frequency of variants per chromosome, **broken down by sample**
- no filtering is applied, all variants are included

```{r}
chromosomes <- sort(unique(df$CHROM))
```

```{r, results = 'asis'}
for (chrom in chromosomes) {
  cat("\n\n#### ", chrom, " {-}\n\n")
  plot_freq_bysample <- df %>%
    filter(CHROM == chrom) %>%
    ggplot(aes(x = POS / 1000, y = QUAL, color = FILE)) +
    geom_point() +
    labs(
      x = "position (kbp)", y = "quality score",
      title = paste0("Variant frequency, chromosome: ", chrom),
      subtitle = "Color: sample"
    ) +
    scale_fill_manual(values = custom_range(length(list_variants), FALSE)) +
    scale_color_manual(values = custom_range(length(list_variants), FALSE)) +
    custom_theme() +
    theme(legend.position = "bottom")
  print(plot_freq_bysample)
  save_plot(plot_freq_bysample, index = chrom)
}
```

### {-}

### Variant frequency by type {.tabset}

- this figure shows the frequency of variants per chromosome, **broken down by type**
- no filtering is applied, all variants are included

```{r, results = 'asis'}
for (chrom in chromosomes) {
  cat("\n\n#### ", chrom, " {-}\n\n")
  plot_freq_bytype <- df %>%
    filter(CHROM == chrom) %>%
    ggplot(aes(x = POS / 1000, y = QUAL, color = VARTYPE)) +
    geom_point() +
    labs(
      x = "position (kbp)", y = "quality score",
      title = paste0("Variant frequency, chromosome: ", chrom),
      subtitle = "Color: variant type"
    ) +
    scale_fill_manual(values = custom_range(length(unique(df$VARTYPE)), FALSE)) +
    scale_color_manual(values = custom_range(length(unique(df$VARTYPE)), FALSE)) +
    custom_theme() +
    theme(legend.position = "bottom")
  print(plot_freq_bytype)
  save_plot(plot_freq_bytype, index = chrom)
}
```

### {-}

### Variant frequency as bar chart {.tabset}

- this figure shows the frequency of **variants per chromosome as a bar chart**
- variants are **broken down by type**
- no filtering is applied, all variants are included

```{r, results = 'asis'}
for (chrom in chromosomes) {
  cat("\n\n#### ", chrom, " {-}\n\n")
  plot_freq_barchart <- df %>%
    filter(CHROM == chrom) %>%
    group_by(FILE, VARTYPE) %>%
    count() %>%
    ggplot(aes(x = n, y = FILE, fill = VARTYPE)) +
    geom_col(color = "white") +
    labs(
      x = "count", y = "",
      title = paste0("Variant frequency, per sample, chromosome: ", chrom),
      subtitle = "Color: variant type"
    ) +
    scale_fill_manual(values = custom_range(length(unique(df$VARTYPE)), FALSE)) +
    scale_color_manual(values = custom_range(length(unique(df$VARTYPE)), FALSE)) +
    custom_theme() +
    theme(legend.position = "bottom")
  print(plot_freq_barchart)
  save_plot(plot_freq_barchart, index = chrom)
}
```

### {-}

### Variant co-occurence, overall similarity

- the following figure shows a **PCA plot** generated from the frequency of the same variants over multiple samples
- samples that show a similar distribution of variants are located closer to each other
- note: filtering is applied such that **only variants with at least N occurences** are included
- the default is *N = 2*, which means that the variant must be present in at least two samples
- can be adjusted using the config parameter `report` > `minimum_variant_count` (current value: `r vcf_count`)
  
```{r}
df_filtered <- df %>%
  inner_join(df_consensus, by = join_by(CHROM, POS, REF, ALT))

no_var_message <- paste0(
  "No variants detected that appear more than ", vcf_count, " times"
)

if (nrow(df_filtered)) {
  df_variants <- df_filtered %>%
    mutate(variant = paste0(CHROM, " [", POS, "] (", REF, "->", ALT, ")")) %>%
    dplyr::select(FILE, variant) %>%
    group_by(variant) %>%
    summarize(files = list(FILE), .groups = "drop")

  df_pca <- df_filtered %>%
    mutate(variant = paste0(CHROM, " [", POS, "] (", REF, "->", ALT, ")")) %>%
    group_by(variant) %>%
    mutate(n = seq_along(FILE)) %>%
    dplyr::select(FILE, variant, n) %>%
    ungroup() %>%
    pivot_wider(names_from = "FILE", values_from = "n") %>%
    mutate(across(everything(), ~ ifelse(is.na(.x), 0, 1))) %>%
    dplyr::select(-variant) %>%
    as.matrix() %>%
    t() %>%
    prcomp() %>%
    {
      .$x
    } %>%
    as_tibble(rownames = "sample")

  for (pc in c("PC2", "PC3")) {
    if (!pc %in% colnames(df_pca)) {
      df_pca[[pc]] <- 0
    }
  }

  plot_pca <- df_pca %>%
    ggplot(aes(
      x = PC1, y = PC2, size = PC3, color = sample,
      label = sample
    )) +
    geom_point() +
    geom_text_repel(size = 3, point.padding = 10, show.legend = FALSE) +
    labs(
      title = "PCA based on presence/absence of identical variants",
      subtitle = "point size encodes PC3, color encodes condition"
    ) +
    custom_theme(aspect.ratio = 1, legend.position = "bottom") +
    guides(color = guide_legend(ncol = 3), size = guide_legend(ncol = 3)) +
    scale_color_manual(values = rep_len(
      custom_colors,
      length.out = nrow(df_pca)
    ))
} else {
  print(no_var_message)
}
```

```{r, fig.width = figwidth, fig.height = figheight*1.5}
if (nrow(df_filtered)) {
  print(plot_pca)
  save_plot(plot_pca, height = figheight * 1.5)
}
```

### Variant co-occurence, frequency

- the following figure contains the same information as in the previous figure,
- but variant co-occurence is summarized in an **upset plot** instead

```{r}
if (nrow(df_filtered)) {
  df_upset <- df_filtered %>%
    mutate(variant = paste0(CHROM, " [", POS, "] (", REF, "->", ALT, ")")) %>%
    dplyr::select(FILE, variant) %>%
    group_by(variant) %>%
    summarize(samples = list(FILE), .groups = "drop")

  plot_upset <- df_upset %>%
    mutate(samples = map(samples, ~ sort(.x))) %>%
    ggplot(aes(x = samples)) +
    geom_bar(position = "stack", fill = custom_colors[7]) +
    scale_x_upset(
      order_by = "degree",
      sets = sort(unique(unlist(df_upset$samples)))
    ) +
    geom_text(
      position = "stack", stat = "count",
      aes(label = after_stat(count)),
      vjust = -0.5, size = 3,
      color = grey(0.3)
    ) +
    labs(
      x = "", y = "count",
      title = "",
      subtitle = ""
    ) +
    custom_theme(legend.position = "right") +
    theme_combmatrix(combmatrix.panel.line.size = 0)
} else {
  print(no_var_message)
}
```

```{r, fig.width = figwidth, fig.height = figheight}
if (nrow(df_filtered)) {
  print(plot_upset)
  save_plot(plot_upset)
}
```

### Variant co-occurence, heat maps {.tabset}

- the following figure shows a heat map of variant occurence per sample
- note: filtering is applied such that **only variants with at least N occurences** are included
- the default is *N = 2*, which means that the variant must be present in at least two samples
- can be adjusted using the config parameter `report` > `minimum_variant_count` (current value: `r vcf_count`)
- heat maps are restricted to a maximum number of 100 variants

```{r, results = 'asis', fig.width = figwidth, fig.height = 12}
for (effect_type in unique(df_filtered$VARTYPE)) {
  if (!is.null(effect_type)) {
    cat("\n\n#### ", effect_type, " {-}\n\n")

    df_vartype <- df_filtered %>%
      filter(VARTYPE == effect_type) %>%
      mutate(FILE = factor(FILE, levels = unique(df$FILE))) %>%
      mutate(variant = paste0(CHROM, " [", POS, "] (", REF, "->", ALT, ")") %>%
        str_trunc(width = 20, side = "right", ellipsis = "…") %>%
        str_pad(width = 20, side = "right", pad = " ")) %>%
      # include also empty tiles for missing values in certain samples
      tidyr::complete(
        FILE,
        variant
      ) %>%
      # restrict to top 100 variants by score
      group_by(variant) %>%
      mutate(mean_qual = mean(QUAL, na.rm = TRUE)) %>%
      ungroup() %>%
      arrange(desc(mean_qual)) %>%
      group_by(FILE) %>%
      dplyr::slice(1:100)

    hm_height <- length(unique(df_vartype$variant))
    hm_width <- length(unique(df_vartype$FILE))

    heatmap <- df_vartype %>%
      ggplot(aes(x = FILE, y = variant, fill = QUAL)) +
      geom_tile() +
      labs(
        x = "", y = "",
        title = paste0("Variant co-occurence, for ", effect_type),
        subtitle = "Color: quality score"
      ) +
      scale_fill_gradientn(colors = custom_range(5, TRUE), na.value = "transparent") +
      custom_theme() +
      theme(
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 6),
        axis.text.y = element_text(size = 6)
      ) +
      coord_fixed(ratio = 0.95)

    print(heatmap)
    save_plot(
      heatmap,
      index = effect_type,
      height = max(c(4, hm_height * 0.25)),
      width = max(c(8, hm_width * 0.25))
    )
  }
}
```

### Variant co-occurence, most affected genes

- this figure shows a bar chart of the **most affected genes** by number of variants
- to qualify, a gene must have **at least two different variants annotated**
- note: for this depiction it does not matter whether variants were found in the same or other samples
- the figure is restricted to the **top 20 genes** with two or more variants

```{r, fig.width = figwidth, fig.height = figheight}
if (nrow(df_filtered)) {
  plot_genes_barchart <- df_filtered %>%
    filter(GENE != "") %>%
    dplyr::select(GENE, CHROM, POS, REF, ALT, VARTYPE) %>%
    distinct() %>%
    group_by(GENE, VARTYPE) %>%
    count() %>%
    group_by(GENE) %>%
    mutate(n_total = sum(n)) %>%
    ungroup() %>%
    arrange(desc(n_total)) %>%
    filter(GENE %in% unique(GENE)[1:20]) %>%
    ggplot(aes(x = n, y = GENE, fill = VARTYPE)) +
    geom_col(color = "white") +
    labs(
      x = "count", y = "",
      title = "Variant frequency, per gene",
      subtitle = "Color: variant type"
    ) +
    scale_fill_manual(values = custom_range(length(unique(df$VARTYPE)), FALSE)) +
    scale_color_manual(values = custom_range(length(unique(df$VARTYPE)), FALSE)) +
    custom_theme()
} else {
  print(no_var_message)
}
```

```{r, fig.width = figwidth, fig.height = figheight}
if (nrow(df_filtered)) {
  print(plot_genes_barchart)
  save_plot(plot_genes_barchart)
}
```

## Variant location and effect

- variants are displayed in decreasing order of impact on target protein
- **intergenic variants are not included for display**, can be looked up in the ouput tables
- note: filtering is applied such that **only variants with at least N occurences** are included
- the default is *N = 2*, which means that the variant must be present in at least two samples
- can be adjusted using the config parameter `report` > `minimum_variant_count` (current value: `r vcf_count`)
- only the **top 10 variants by quality score** are shown to keep the report concise

```{r}
filter_variants <- function(df, vartype, topn) {
  cols_select <- c(
    "CHROM", "POS", "REF", "ALT", "VARTYPE", "IMPACT",
    "GENE", "AA_CHANGE", "CODON_CHANGE", "INDEL"
  )
  if ("VARTYPE" %in% colnames(df)) {
    return(
      df %>%
        filter(VARTYPE == vartype) %>%
        group_by(across(any_of(cols_select))) %>%
        summarize(samples = list(FILE), QUAL = mean(QUAL), .groups = "drop") %>%
        dplyr::arrange(desc(QUAL)) %>%
        dplyr::slice(1:topn)
    )
  } else {
    return(tibble())
  }
}

plot_genome <- function(df, df_mut, tx_width = 1) {
  tx_window <- df_mut[["POS"]] + c(-5000, 5000)
  df %>%
    filter(seqnames == df_mut$CHROM[1]) %>%
    filter(end >= tx_window[1] & start <= tx_window[2]) %>%
    mutate(
      xstart = ifelse(strand == "+", start, end),
      xend = ifelse(strand == "+", end, start),
      y = ifelse(strand == "+", 0.25, -0.25)
    ) %>%
    ggplot(aes(color = tx_name, fill = tx_name, label = tx_name)) +
    geom_segment(
      x = tx_window[1], y = 0, xend = tx_window[2], yend = 0,
      linewidth = 2, lineend = "square", linejoin = "mitre", color = grey(0.7)
    ) +
    geom_segment(aes(x = 0, y = 0, xend = tx_width, yend = 0),
      linewidth = 1.33, lineend = "square", linejoin = "mitre", arrow = arrow(
        angle = 30, length = unit(0.02, "inches"),
        type = "closed"
      )
    ) +
    geom_segment(
      aes(x = xstart, y = y, xend = xend, yend = y),
      linewidth = 1.5, lineend = "square", linejoin = "mitre", arrow = arrow(
        angle = 30, length = unit(0.01, "inches"),
        type = "closed"
      )
    ) +
    geom_text(
      aes(
        x = ifelse(strand == "+", xstart + (xend - xstart) / 2, xend - (xend - xstart) / 2),
        y = ifelse(strand == "+", 0.35, -0.35)
      ),
      size = 2.0, hjust = 0.5
    ) +
    geom_segment(
      data = mutate(df_mut, tx_name = NA),
      aes(x = POS - 5, y = 0, xend = POS + 5, yend = 0),
      linewidth = 2, lineend = "square", linejoin = "mitre", color = "red"
    ) +
    geom_label(
      data = mutate(df_mut, tx_name = NA),
      aes(x = POS, y = 0, label = paste(c(CHROM, POS, IMPACT, GENE, CODON_CHANGE), collapse = ", ")),
      size = 2.5, fill = "white", color = "red", hjust = -0.05,
    ) +
    coord_cartesian(ylim = c(-0.5, 0.5), xlim = tx_window, expand = FALSE) +
    labs(
      x = "position [nt]", y = "",
      title = with(df_mut, paste0(
        "CHR: ", CHROM, ", pos: ", POS, ", impact: ", IMPACT, ", gene: ", GENE,
        ", codon: ", CODON_CHANGE
      )),
      subtitle = paste0("Variant present in: ", paste(df_mut[["samples"]], collapse = ", "))
    ) +
    custom_theme(legend.position = "none") +
    scale_color_manual(values = rep_len(
      custom_colors,
      length.out = nrow(df_tx)
    )) +
    theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())
}

df_tx <- annotated_transcripts %>%
  as_tibble()
topn <- 10
```

### Stop lost

```{r, fig.width = figwidth, fig.height = 2.0}
df_plotting <- filter_variants(df_filtered, "stop_lost", topn)

if (nrow(df_plotting)) {
  for (i in 1:min(c(topn, nrow(df_plotting)))) {
    plot_genome_stop_lost <- plot_genome(df_tx, df_plotting[i, ])
    print(plot_genome_stop_lost)
  }
} else {
  print(no_var_message)
}
```

### Stop gained

```{r, fig.width = figwidth, fig.height = 2.0}
df_plotting <- filter_variants(df_filtered, "stop_gained", topn)

if (nrow(df_plotting)) {
  for (i in 1:min(c(topn, nrow(df_plotting)))) {
    plot_genome_stop_gained <- plot_genome(df_tx, df_plotting[i, ])
    print(plot_genome_stop_gained)
  }
} else {
  print(no_var_message)
}
```

### Frameshift variant

```{r, fig.width = figwidth, fig.height = 2.0}
df_plotting <- filter_variants(df_filtered, "frameshift_variant", topn)

if (nrow(df_plotting)) {
  for (i in 1:min(c(topn, nrow(df_plotting)))) {
    plot_genome_frameshift_variant <- plot_genome(df_tx, df_plotting[i, ])
    print(plot_genome_frameshift_variant)
  }
} else {
  print(no_var_message)
}
```

### Missense variant

```{r, fig.width = figwidth, fig.height = 2.0}
df_plotting <- filter_variants(df_filtered, "missense_variant", topn)

if (nrow(df_plotting)) {
  for (i in 1:min(c(topn, nrow(df_plotting)))) {
    plot_genome_missense_variant <- plot_genome(df_tx, df_plotting[i, ])
    print(plot_genome_missense_variant)
  }
} else {
  print(no_var_message)
}
```

# About this report

## Workflow

This report was automatically generated by the [snakemake-simple-mapping](https://github.com/MPUSP/snakemake-simple-mapping) workflow.

For issues, bugs, and feature requests please use the workflow's [github page](https://github.com/MPUSP/snakemake-simple-mapping/issues).

For all other feedback, contact the author(s):

- Michael Jahn, PhD (author and maintainer): jahn@mpusp.mpg.de

The workflow is developed at the [Max Planck Unit for the Science of Pathogens](https://www.mpusp.mpg.de/), Berlin, Germany.

## Configuration

The pipeline configuration file is available at: `config/config.yml`.

```{r}
list_config <- read_lines("config/config.yml")
list_config <- gsub("\"", "", list_config)
print(list_config)
```

## Data accessability

The following resources were used to generate this report:

- VCF input for each sample: `r list_variants`
- VCF consensus between all samples: `r list_consensus`

## Session Info

Link to the R markdown source that was used to generate this report:

<a download="report.Rmd" href="`r base64enc::dataURI(file = params$rmd, mime = 'text/rmd', encoding = 'base64')`">R Markdown source file</a>

Session info (R base and package versions):

```{r}
sessionInfo()
```

```{r}
messages <- append(messages, "finished writing HTML report successfully")
write_lines(
  paste0("HTML_REPORT: ", messages),
  file = "results/report/report.log"
)
```
